---
title: "Customer Segmentation using K-means Clustering in R"
author: "**Akash Sukhavasi**"
date: "December 12th, 2020"
output: html_notebook
---

Final Markdown of Project

# Installing Required Packages -> "factoextra","NbClust","ggplot2","plotrix"
```{r}
pkgs <- c("factoextra","NbClust","ggplot2","plotrix")
install.packages(pkgs)
```

# Reading Customer Data-set into a Variable & Displaying Customer
```{r}
df = read.csv("Customer Data.csv")
df
```

# Cleaning Customer Data (Omitting Null Values)
```{r}
df <- na.omit(df)
```

# Displaying properties/attributes of Data
```{r}
names(df)
```

# Displaying First 10 Rows
```{r}
head(df)
```

# Structure of Data
```{r}
str(df)
```

# Making a Duplicate of Data & Presenting a Summary
```{r}
customer_data = df
summary(customer_data)
```
# Standard Deviation of Age
```{r}
sd(customer_data$Age)
```
# Standard Devation of Income
```{r}
summary(customer_data$Annual.Income..k..)
```
# Standard Deviation of Spending Score
```{r}
sd(customer_data$Spending.Score..1.100.)
```

# Gender Comparision
Bar Plot for Gender Comparison
```{r}
a=table(customer_data$Gender)
barplot(a,main="Using BarPlot to display Gender Comparision",
        ylab="Count",
        xlab="Gender",
        col=rainbow(2),
        legend=rownames(a))
```

Pie Chart describing the percentages of Male and Female Gender
```{r}
pct=round(a/sum(a)*100)
lbs=paste(c("Female","Male")," ",pct,"%",sep=" ")
library(plotrix)
pie3D(a,labels=lbs,
      main="Pie Chart Depicting Ratio of Female and Male")
```

Box Plot for descriptive Analysis of Spending Score
```{r}
boxplot(customer_data$Spending.Score..1.100.,
        horizontal=TRUE,
        col="#990000",
        main="BoxPlot for Descriptive Analysis of Spending Score")
```

# Summary Of Spending Score
```{r}
summary(customer_data$Spending.Score..1.100.)
```

# Visualization of Age and Gender together
```{r}
library(ggplot2)
bold_axis <- element_text(face = "bold", color = "black", size = 20)
axis_text <- element_text(face = "bold", size = 14)
# bold axis is used for the text characteristics of headings of the graphs
# axis_text is used for the text characteristics of labels of the graphs
```

# Histogram of Customer Age
```{r}
Plotage <- ggplot(customer_data,aes(x=Age))
Plotage + geom_histogram(fill="purple", alpha = 0.7,bins=30) + theme(axis.text = axis_text) + theme(title = bold_axis) + ggtitle("Histogram of Customer Age")
```

# Visualization of customers by age and gender
```{r}
Plotagegender <- ggplot(customer_data, aes(x = Age))

Plotagegender + geom_histogram(aes(fill = factor(Gender)),bins=30) + ggtitle("Histogram of Customers by Age (Colored by Gender)") + theme(axis.text = axis_text) + theme(title = bold_axis)
```

# K-Means Clustering

K-means is a centroid-based algorithm, or a distance-based algorithm, where we calculate the distances to assign a point to a cluster. In K-Means, each cluster is associated with a centroid.

There are three methods to find the optimal number of clusters :
1.Elbow Method
2.Silhouette Method
3.Gap-statistic Method

we need libraries like ggplot2,dplyr,cluster,factoextra for implementing k-means clustering

```{r}
library(ggplot2)
library(dplyr)
library(cluster)
library(factoextra)
```

For simplicity reasons , we are renaming the attribute names to cust_id,gender,age,annual_income_000,spend_score
```{r}
colnames(customer_data) <- c("cust_id","gender","age","annual_income_000","spend_score")
```

# 1. Elbow Method
In cluster analysis, the elbow method is a heuristic used in determining the number of clusters in a data set. The method consists of plotting the explained variation as a function of the number of clusters, and picking the elbow of the curve as the number of clusters to use. 

Here we are performing the elbow method on spending_score and annual_income.
```{r}
kk <- 1:10
wss <- c()
for (i in kk){
  cluster_temp <- kmeans(customer_data[,c(4,5)], centers = i, nstart = 25)
  wss <- c(wss, cluster_temp$tot.withinss)
}
plot(wss, type = "b")
```

# 2. Average Silhouette Method
```{r}
fviz_nbclust(customer_data[,c(4,5)], kmeans, method = "silhouette", k.max = 10) + theme_minimal() + ggtitle("The Silhouette Plot")
```

# 3. Gap-Statistic Method
```{r}
set.seed(123)
gap_stat <- clusGap(customer_data[,c(4,5)], FUN = kmeans, nstart = 25, K.max = 10, B = 50)
fviz_gap_stat(gap_stat) + theme_minimal() + ggtitle("Gap Statistic Method")
```

# Gap - Statistic Method Explanation
```{r}
set.seed(123)
gap_stat <- clusGap(customer_data[,c(4,5)], FUN = kmeans, nstart = 25, K.max = 10, B = 50)
print(gap_stat, method = "firstmax")
```

From the above methods the optimal number of clustering can be done with either k=5 or k=6
# K = 5
Clustering According to Spending Score and Annual Income:
considering the number of optimal number of clusters = 5

```{r}
set.seed(10)
cluster_5 <- kmeans(customer_data[,c(4,5)], centers = 5, nstart = 25)
cluster_5
```

As we see in the Clustering Vector all the 200 points in the data-set have been classified into 5 different clusters.

Now we are modifying the data-set by adding the cluster number to each corresponding point using mutate.
```{r}
library(dplyr)
library(magrittr)
customer_data <- customer_data %>% mutate(cluster5 = cluster_5$cluster)
```

Printing the data-set after adding the cluster number
```{r}
customer_data
```

# Printing a Scatter Plot for the above data (K = 5)
```{r}
library(ggplot2)
ggplot(customer_data, aes(x = annual_income_000, y = spend_score, colour = factor(cluster5))) +
  geom_point() +
  scale_color_discrete(name = "Clusters") +
  xlab("Annual income of customers (thousands)") +
  ylab("Spending score")
```
The conclusions from the above graph are :

cluster 1 : High Annual Income and High Spending Score
Cluster 2 : Low Annual Income and High Spending Score
Cluster 3 : Average Annual Income and Average Spending Score
Cluster 4 : High Annual Income and Low spending score
Cluster 5 : Low Annual Income and Low spending score

Till now we considered only the attributes spending score and annual income

# K = 6
```{r}
k6<-kmeans(customer_data[,3:5],6,iter.max=100,nstart=100)
k6
```

# Plotting Customer Age with Spending Score (K = 6)
```{r}
ggplot(df, aes(x =Spending.Score..1.100., y =Age)) + 
  geom_point(aes(color = as.factor(k6$cluster))) +
  scale_color_discrete(name=" ",
                      breaks=c("1", "2", "3", "4", "5","6"),
                      labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5","Cluster 6")) +
  ggtitle("Segments of Mall Customers", subtitle = "Using K-means Clustering")
```

# Principal Component Analysis

As there are three attributes for the execution: Age,Spending Score,Annual Income.

We will be getting a three dimensional point (x,y,z)
Principal component analysis is the process of flattening the data set so that we can convert the three dimensional point into a two dimensional point.

We will be getting 3 rotation points PC1,PC2,PC3
where the variation among these is in the following order : PC1>PC2>PC3
```{r}
pcclust=prcomp(customer_data[,3:5])
pcclust
```

# Summary of Principal Component Analysis
```{r}
summary(pcclust)
```

as PC1 has the highest variation

```{r}
pcclust$rotation[,1:2]
```


```{r}
kCols=function(vec){cols=rainbow (length (unique (vec)))
return (cols[as.numeric(as.factor(vec))])}

digCluster<-k6$cluster; dignm<-as.character(digCluster); # K-means clusters

plot(pcclust$x[,1:2], col = kCols(digCluster),pch = 19,xlab = "K-means",ylab= "Classes")
legend("bottomleft",unique(dignm),fill=unique(kCols(digCluster)))

```